---
title: "Analysis of Biowulf Waiting Times"
date: July 27 2021
author: Gianluca Pegoraro
output: html_document
---


### Analysis Setup
Load required packages.
```{r load-packages}
library(tidyverse)
library(lubridate)
library(ggthemes)
```

```{r set-knitr-options}
knitr::opts_chunk$set(
  fig.path = 'output/',
  message = FALSE,
  warning = FALSE
  )
```

Set the palette and the running theme for ggplot2.
```{r set-theme-palette}
theme_set(theme_bw())
theme_update(axis.text.x = element_text(
angle = -45,
hjust = 0,
vjust = 0.5
))
```

Read Biowulf logs from Susan Chacko. 
```{r read_data}
data <- read_tsv("data/hitif_jobs") %>% 
  rename(q_time = queued_time) %>%
  mutate(q_time = na_if(q_time, "-")) %>%
  separate(gres, 
           into = c("gpu_t", "lscratch_t"), 
           sep = ",") %>%
  separate(gpu_t, 
           into = c("gpu_t", "gpu_type", "gpu_n"), 
           sep = ":",
           convert = TRUE) %>%
  separate(lscratch_t, 
           into = c("lscratch_t", "lscratch_g"),
           sep = ":",
           convert = TRUE) %>%
  select(user, partition, submit_time, 
         gpu_type, gpu_n, lscratch_g,
         q_time)

glimpse(data)
```
Tidy up waiting time format. 
```{r transform-durations}
q_time <- data %>% pull(q_time)

q_time_trans <- q_time %>% 
  str_replace("^(\\d:\\d{2})", "0\\1") %>% 
  str_replace("^(\\d{2}:\\d{2})$", "00:\\1") %>%
  str_replace("^(\\d{2}:\\d{2}:\\d{2})", "1970-01-01 \\1") %>%
  str_replace("^\\d-(\\d{2}:\\d{2}:\\d{2})", "1970-01-02 \\1") %>% 
  ymd_hms()

(q_secs <- as.integer(q_time_trans - origin))
```
Replace queueing times in the data table.
```{r transform-q-time}
queue_durations <- data %>% 
  select(-q_time) %>%
  mutate(q_secs = q_secs)

glimpse(queue_durations)
```

Percentage of requests with no queue time value (incorrect or too fast).
```{r na-stats}
queue_durations %>% 
  summarise(perc_na = 100*mean(is.na(q_secs))) 
```
Percentage of GPU requests by user. 
```{r users-stats}
tot <- nrow(queue_durations)

queue_durations %>%
  group_by(user) %>%
  summarise(user_perc = 100*(n()/tot)) %>%
  arrange(desc(user_perc))
```

Percentage of GPU requests by GPU type.
```{r gpu_type_perc}
queue_durations %>%
  group_by(gpu_type) %>%
  summarise(gpu_type_perc = 100*(n()/tot)) %>%
  arrange(desc(gpu_type_perc))
```

```{r plot-hist-durations, echo = FALSE}
queue_durations %>%
  ggplot(aes(x = gpu_type, 
             y = q_secs/3600,
             color = user)) +
  geom_jitter(position = position_jitter(width = 0.3)) +
  scale_y_continuous(breaks = seq(0, 48, 6)) +
  scale_color_tableau() +
  xlab("GPU Type") +
  ylab("Queue Waiting Time (Hrs)")
```
```{r by-gpu-number, echo = FALSE}
queue_durations %>%
  ggplot(aes(x = gpu_type, 
             y = q_secs/3600,
             color = factor(gpu_n))) +
  geom_jitter(position = position_jitter(width = 0.3)) +
  scale_y_continuous(breaks = seq(0, 48, 6)) +
  scale_color_tableau(name = "# of GPU") +
  xlab("GPU Type") +
  ylab("Queue Waiting Time (Hrs)")
```
```{r by-user, echo = FALSE}
queue_durations %>%
  ggplot(aes(x = gpu_type, 
             y = q_secs/3600,
             color = user)) +
  geom_jitter(position = position_jitter(width = 0.3)) +
  geom_hline(yintercept = 2,
             color = "grey30",
             linetype = "longdash") +
  scale_y_continuous(breaks = seq(0, 6, 0.5)) +
  scale_color_tableau() +
  coord_cartesian(ylim = c(0, 6)) +
  xlab("GPU Type") +
  ylab("Queue Waiting Time (Min)")
```

```{r by-gpu-type, echo = FALSE}
queue_durations %>%
  ggplot(aes(x = submit_time, 
             y = q_secs/3600,
             color = gpu_type)) +
  geom_point() +
  geom_hline(yintercept = 2,
             color = "grey30",
             linetype = "longdash") +
  scale_x_datetime() +
  scale_y_continuous(breaks = seq(0, 48, 6)) +
  scale_color_tableau(name = "GPU Type") +
  xlab("Date") +
  ylab("Waiting Time (Hrs)")
```
Percentage of requests with a waiting time more than a certain threshold. 
```{r perc-long}
queue_durations %>%
  filter(!is.na(q_secs)) %>%
  group_by(gpu_type) %>%
  summarise(n = n(),
            perc_long_1hrs = 100*mean(q_secs >= 3600),
            perc_long_2hrs = 100*mean(q_secs >= 7200),
            perc_long_3hrs = 100*mean(q_secs >= 10800))
```

